BlueTracer: \\ a Robust API Tracer for Evasive Malware

### 2018-07-18 Msc-Slides/main.tex
Daniele's corrections applied

    %Two main types:
    \medskip
    \begin{itemize}
        \item \textcolor{sapienza}{\textbf{Static Analysis:}}\\
        involves the inspection of the different data and code sections of a binary
        \item \textcolor{sapienza}{\textbf{Dynamic Analysis:}}\\
        the malware sample is executed and the actions it performs on the environment are observed
    \end{itemize}
    \vspace{0.5cm}

@@

\begin{itemize}
\item Functions can abstract
implementation details providing
a semantically richer representation of some functionality
\item The abstractions embodied by \textbf{system calls} and \textbf{library calls} \\ can be used to grasp the visible behavior of a malicious sample 
\end{itemize}
\medskip
\textcolor{sapienza}{\textbf{Example}}:\\
\bigskip

@@

Available API tracing tools have \textbf{limited logging capabilities}
%\begin{itemize}
%\item[--] The amount of recorded information is too little
%\end{itemize}
\medskip
	\begin{block}{}
\begin{lstlisting}[basicstyle=\ttfamily\large,xleftmargin=50pt]
<@\textcolor{red}{record\_before(libcall\_name, arg1)}@> 
retval = libcall(arg1, &arg2)
<@\textcolor{red}{record\_after(retval, *arg2)}@> 
\end{lstlisting}
\end{block}	
\medskip
\begin{itemize}
\item Access to a source of calls related information is required
\begin{itemize}
\item[--] Prototype (number of arguments, data types, input/output)
%\item[--] Extraction
\end{itemize}
\medskip
\item \textbf{Challenge:} heterogeneity of Windows libraries used in malware and lack of well-structured documentation for their prototypes
\end{itemize}

@@

    \begin{figure}
    	\vspace{-0.2cm}
    	\hspace*{-1cm}
        \includegraphics[width=12cm]{image/MainImage1.pdf}
    \end{figure}

@@

\textcolor{sapienza}{\textbf{The tool possesses a remarkable logging power:}}
\begin{itemize}
%\item Integration of reliable external sources \\ (\textbf{Dr. %Memory}
%and \textbf{CISCO PyREBox})
%\item Logging operations are challenging 
%by the heterogeneity of Windows libraries used in malware and the lack of well-structured documentation for their prototypes
%\begin{itemize}
\item Integration of reliable external sources \\ (\textbf{Dr. Memory}
and \textbf{CISCO PyREBox})
%\end{itemize} 
%\end{itemize}

@@

Only calls made directly by the sample are recorded through the use of context-sensitive introspection
\begin{itemize}
\item Stack pointer inspection
\end{itemize}    
\medskip
	    \begin{figure}
    	\vspace{-0.3cm}
    	%\hspace*{1cm}
        \includegraphics[width=10cm]{image/MainImageTree.pdf}
    \end{figure}

@@

\begin{frame}
    \frametitle{BlueTracer's architecture}

@@

\subsection{Al-Khaser}
\begin{frame}
    \frametitle{Validation (1) with Al-Khaser}
    \begin{beamerboxesrounded}[shadow=true]{}
	\textbf{Al-Khaser} is an open-source application which performs common checks employed by malware families to determine if they are being
executed in an analysis environment.
\end{beamerboxesrounded}
\medskip
Many implemented techniques:
\begin{itemize}
\item \textbf{Anti-Debugging}
\item \textbf{Timing-based}
\item \textbf{Human Interaction Detection}
\item \textbf{Anti-Virtualization}
\item \textbf{Detection of Analysis Tools}
\end{itemize}
\medskip
\begin{beamerboxesrounded}[shadow=true]{}
BlueTracer: 
\begin{itemize}
\item remained undetected thanks to its integration with BluePill
\item managed to track all the checks
\end{itemize}
\end{beamerboxesrounded}

@@

    \begin{table}[h]
\vspace*{-0.1cm}
\begin{center}
\resizebox{9.2cm}{!}{
 \begin{tabular}{||c c c||}
 \hline
 \textbf{ID} & \textbf{MD5} & \textbf{Name} \\ 
 \hline\hline
 1 & 0af4ef5069f47a371a0caf22ae2006a6 & \textit{banker} \\ 
 \hline
 2 & 9437eabf2fe5d32101e3fbf9f6027880 & \textit{dropper} \\
 \hline
 3 & cbdda646a20d95f078393506ecdc0796 & \textit{trojan}\\
 \hline
 4 & cfdd16225e67471f5ef54cab9b3a5558 & Olympic\\
 \hline
 5 & ef694b89ad7addb9a16bb6f26f1efaf7 & Malicious CCleaner\\ 
 \hline
\end{tabular}
}
\end{center}
\end{table}
\vspace{-0.1cm}
Evaluation was done manually and is a time-consuming process:
\begin{itemize}
\item Check if logs are congruous with Joe Security reports
\item Process Monitor as ground truth for system activity
\end{itemize}
\medskip
\begin{beamerboxesrounded}[shadow=true]{}
The
logs collected by BlueTracer reveal behaviors consistent with the analysis reports
authored by Joe Security
\end{beamerboxesrounded}

@@

\begin{lstlisting}
~~1116~~ 138 kernel32.dll!CopyFileA
138 	arg 0: c:\Users\Simuset\Desktop\sample1.exe 
(name=lpExistingFileName, type=char*, size=0x1)
138 	arg 1: <@\alert{C:\textbackslash Windows\textbackslash system32\textbackslash $\dagger$\textbackslash ffpb6966.exe}@> 
(name=lpNewFileName, type=char*, size=0x1)
138 	arg 2: 0x0 (name=bFailIfExists, type=(long/int), size=0x4)
138    executed kernel32.dll!CopyFileA =>
138 	retval: 0x1 (name=Return value, type=(long/int), size=0x4)
\end{lstlisting}

@@

	\textbf{\textcolor{sapienza}{Contribution:}} \\
	\smallskip	
	\tab Design and implementation of \textbf{BlueTracer},
a robust library and system call tracer for Windows programs specialized in evasive
malware.
	\\\bigskip
	\textbf{\textcolor{sapienza}{Future Developments:}} \\
	\begin{itemize}
	\item Automatic methodology for large-scale evaluation 
	\item Improvement of logging capabilities
	\item Usage of log filtering and aggregation techniques
	\end{itemize}

>>>>

    %Two main types:
    \medskip
    \begin{itemize}
        \item \textcolor{sapienza}{\textbf{Static Analysis:}}\\
        involves the inspection of the different data and code sections of a binary
        \item \textcolor{sapienza}{\textbf{Dynamic Analysis:}}\\
        a malware sample is executed and the actions it performs on the environment are observed
    \end{itemize}
    \vspace{0.5cm}

@@

\begin{itemize}
\item Functions can abstract
implementation details providing
a semantically richer representation of some functionality
\item The abstractions embodied by \textbf{system calls} and \\ \textbf{library functions} can be used to grasp the visible behavior of a malicious sample 
\end{itemize}
\medskip
\textcolor{sapienza}{\textbf{Example}}:\\
\bigskip

@@

Available API tracing tools have \textbf{limited logging capabilities}
%\begin{itemize}
%\item[--] The amount of recorded information is too little
%\end{itemize}
\medskip
	\begin{block}{}
\begin{lstlisting}[basicstyle=\ttfamily\large,xleftmargin=50pt]
<@\textcolor{red}{record\_before(libcall\_name, arg1)}@> 
retval = libcall(arg1, &arg2)
<@\textcolor{red}{record\_after(retval, *arg2)}@> 
\end{lstlisting}
\end{block}	
\medskip
\begin{itemize}
\item Access to calls related information is required
\begin{itemize}
\item[--] Prototype (number of arguments, data types, input/output)
\item[--] Need to know their run-time locations
\end{itemize}
\medskip
\item \textbf{Challenge:} heterogeneity of Windows libraries used in malware and lack of well-structured documentation for their prototypes
\end{itemize}

@@

    \begin{figure}
    	\vspace{-0.2cm}
    	\hspace*{-0.2cm}
        \includegraphics[width=10cm]{image/MainImage11.pdf}
    \end{figure}

@@

\textcolor{sapienza}{\textbf{The tool possesses a remarkable logging power:}}
\begin{itemize}
%\item Integration of reliable external sources \\ (\textbf{Dr. %Memory}
%and \textbf{CISCO PyREBox})
%\item Logging operations are challenging 
%by the heterogeneity of Windows libraries used in malware and the lack of well-structured documentation for their prototypes
%\begin{itemize}
\item Integration of rich, reliable external sources \\ (\textbf{Dr. Memory}
and \textbf{CISCO PyREBox})
%\end{itemize} 
%\end{itemize}

@@

Only calls made directly by the sample are recorded through the use of context-sensitive introspection
\begin{itemize}
\item Stack pointer and return address analysis
\end{itemize}    
\medskip
	    \begin{figure}
    	\vspace{-0.3cm}
    	%\hspace*{1cm}
        \includegraphics[width=10cm]{image/MainImageTree.pdf}
    \end{figure}

@@

\begin{frame}
    \frametitle{Architecture of BlueTracer}

@@

\subsection{Al-Khaser}
\begin{frame}
    \frametitle{Validation (1): Al-Khaser}
    \begin{beamerboxesrounded}[shadow=true]{}
	\textbf{Al-Khaser} is an open-source application which performs checks \\ popular in malware samples spotted in the wild to determine if they are being
executed in an analysis environment.
\end{beamerboxesrounded}
\medskip
Many implemented techniques:
\begin{itemize}
\item \textbf{Anti-Debugging}
\item \textbf{Timing-based}
\item \textbf{Human Interaction Detection}
\item \textbf{Anti-Virtualization}
\item \textbf{Detection of Analysis Tools}
\end{itemize}
\medskip
\begin{beamerboxesrounded}[shadow=true]{}
BlueTracer: 
\begin{itemize}
\item remained undetected thanks to its integration with BluePill
\item managed to track all arguments used to trigger evasion
\end{itemize}
\end{beamerboxesrounded}

@@

    \begin{table}[h]
\vspace*{-0.1cm}
\begin{center}
\resizebox{9.2cm}{!}{
 \begin{tabular}{||c c c||}
 \hline
 \textbf{ID} & \textbf{MD5} & \textbf{Name} \\ 
 \hline\hline
 1 & 0af4ef5069f47a371a0caf22ae2006a6 & \textit{banker} \\ 
 \hline
 2 & 9437eabf2fe5d32101e3fbf9f6027880 & \textit{dropper} \\
 \hline
 3 & cbdda646a20d95f078393506ecdc0796 & \textit{trojan}\\
 \hline
 4 & cfdd16225e67471f5ef54cab9b3a5558 & Olympic\\
 \hline
 5 & ef694b89ad7addb9a16bb6f26f1efaf7 & Malicious CCleaner\\ 
 \hline
\end{tabular}
}
\end{center}
\end{table}
\vspace{-0.1cm}
Manual validation of all relevant actions performed by each sample:
\vspace{-0.5cm}
\begin{itemize}
\item Check if logs are congruous with Joe Security reports
\item Process Monitor as ground truth for system activity
\end{itemize}
\medskip
\begin{beamerboxesrounded}[shadow=true]{}
The
logs collected by BlueTracer reveal behaviors consistent with the analysis reports
authored by Joe Security
\end{beamerboxesrounded}

@@

\begin{lstlisting}
~~1116~~ 138 kernel32.dll!CopyFileA
138 	arg 0: <@\alert{C:\textbackslash Users\textbackslash Simuset\textbackslash Desktop\textbackslash sample1.exe}@> 
(name=lpExistingFileName, type=char*, size=0x1)
138 	arg 1: <@\alert{C:\textbackslash Windows\textbackslash system32\textbackslash $\dagger$\textbackslash ffpb6966.exe}@> 
(name=lpNewFileName, type=char*, size=0x1)
138 	arg 2: <@\alert{0x0}@> (name=bFailIfExists, type=(long/int), size=0x4)
138    executed kernel32.dll!CopyFileA =>
138 	retval:<@\alert{0x1}@> (name=Return value, type=(long/int), size=0x4)
\end{lstlisting}

@@

	\textbf{\textcolor{sapienza}{Contribution:}} \\
	\smallskip	
	\tab Design and implementation of \textbf{BlueTracer},
a robust library and \tab system call tracer for Windows programs specialized in evasive
\tab malware.
	\\\bigskip
	\textbf{\textcolor{sapienza}{Future Developments:}} \\
	\begin{itemize}
	\item Automatic methodology for large-scale evaluation 
	\item Improvement of logging capabilities
	\item Log filtering and aggregation techniques
	\end{itemize}
